---
title: "p8105_hw6_niz2000"
author: "Nora Zakaria"
date: "11/25/2019"
output: github_document
---

## Load Tidyverse and Class Datasets

```{r}
library(tidyverse)
library(p8105.datasets)
```


# Problem 1

## Load and Tidy the Birthweight Data

```{r warning=FALSE}
#Load and Tidy the Data
birthweight = 
  read_csv(file = "./data/birthweight.csv") %>%
  janitor::clean_names() %>%
  mutate(
    babysex = as.factor(babysex),
      babysex = fct_recode(babysex,
                  "male" = "1", "female" = "2"),
    frace = as.factor(frace),
      frace = fct_recode(frace,
                  "white" = "1", "black" = "2", "asian" = "3", 
                  "Puerto Rican" = "4", "Other" = "8", "Unknown" = "9"),
    malform = as.factor(malform),
      malform = fct_recode(malform,
                  "absent" = "0", "present" = "1"),             
    mrace = as.factor(mrace),
      mrace = fct_recode(mrace,
                  "white" = "1", "black" = "2", "asian" = "3", 
                  "Puerto Rican" = "4", "Other" = "8"))
birthweight

# Check for Missing Values
missing_test = birthweight %>%
  summarise_all((funs(sum(is.na(.))))) 
missing_test
```

The four categorical variables in the dataset, babysex, frace, malform, and mrace, were converted from numeric to factor variables. These factors were also recoded in order for variable codes to be more descriptive. For example in the babysex variable, the factors now display male and female, opposed to 1 and 2. After tidying, the missing_test dataframe was created to detect if there are any missing values in the dataset. As we see in the missing_test outputted table, there are 0 missing values in each of the variable columns.


## Regression Model for Birthweight

In this linear regression model for the birthweight (bwt) outcome, predictors include mother's pre-pregnancy BMI (ppbmi), mother’s weight gain during pregnancy in pounds (wtgain), mother's age at delivery in years (momage), the number of live births prior to this pregnancy (parity), and the average number of cigarettes smoked per day during pregnancy (smoken). These variables were selected based on previous academic literature findings on the factors that underly birthweight. In addition to mother's starting weight, weight gain during preganancy, and age, birth order and behavioral factors such as smoking all contribute to the birthweight of a newborn. 

```{r Fit Linear Model}
fit = lm(bwt ~ ppbmi + wtgain + momage + parity + smoken, data = birthweight)
fit

fit_summary_table = fit %>%
  broom::tidy() %>%
  knitr::kable(digits = 3) 
fit_summary_table
```

The linear model summary table provides the regression coefficients, standard error, test statistic, and p-value for all predictors in our model. All p-values were less than 0.05, apart from the "parity" variable, therefore "parity" is the only predictor in our model that is not statistically significant. Additionally, our regression coefficients for "parity" and "smoken" are negative, therefore demonstrating a negative linear relationship between the predictor and our birthweight outcome.


## Plot of Model Residuals and Fitted Values

```{r Model Plot}
modelr::add_predictions(birthweight, fit)
modelr::add_residuals(birthweight, fit)

birthweight %>%
  modelr::add_residuals(fit) %>%
  modelr::add_predictions(fit) %>%
  ggplot(aes(x = pred, y = resid)) +
    geom_point(aes(color = "viridis"), show.legend = FALSE) +
    labs(
      title = "Model Residuals Against Fitted Values",
      x = "Fitted Values",
      y = "Residuals")  +
    viridis::scale_color_viridis(discrete = TRUE) 
```

The scatterplot of residuals against fitted values is concentrated around fitted values from 2800 to 33000, and residual values from -1000 to 1000.


## Comparing Models

Next, the "fit" model will be compared to two additional models:

* The "length_age_fit" model, including length at birth and gestational age as predictors; and
* The "interaction_terms_fit" model, including head circumference, length at birth, sex, and all interacting terms as predictors.

### Predictors: Length at Birth and Gestational Age

```{r Length_Age_Fit Model}
length_age_fit = lm(bwt ~ blength + gaweeks, data = birthweight)
```

### Predictors: Head Circumference, Length, Sex, and Interacting Terms

```{r Interaction_Terms_Fit Model}
interaction_terms_fit = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = birthweight)
```


Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.


# Problem 2

## Load 2017 Central Park NOAA Weather Data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


